---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---


## Loading libraries
```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
```


## Loading and preprocessing the data

Extract data from zip file into a data folder if it doesn't exists.

```{r, echo=TRUE}
if(!file.exists("./data")) {
  dir.create("./data")
  unzip(zipfile = "./activity.zip", exdir = "./data")
}
```

Load activity data from extracted file.

```{r, echo=TRUE}
activity <- read.csv('./data/activity.csv', stringsAsFactors = FALSE)
```

Explicitly convert the date field.

```{r, echo=TRUE}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
```


## What is mean total number of steps taken per day?


### Calculate total number of steps taken per day    

```{r, echo=TRUE}
activity.per.day <- activity %>% group_by(date) %>% summarize (steps = sum(steps))
```


### Show the total number of steps per day with an histogram

```{r, echo=TRUE, fig.path="./figure/"}
qplot(date, data=activity.per.day, weight=steps, geom="histogram", binwidth=1)
```


### Mean and median of the total number of steps taken per day

```{r, echo=TRUE}
round(mean(activity.per.day$steps, na.rm = T))
round(median(activity.per.day$steps, na.rm = T))
```

## What is the average daily activity pattern?

### Average of steps taken across all 5 minutes inverval

Calculate the average steps for each 5 minutes interval.

```{r, echo=TRUE}
activity.per.5mins <- activity %>% 
  group_by(interval) %>% 
  summarize (steps.avg = round(mean(steps, na.rm = T)))
```

Plot the result.

```{r, echo=TRUE, fig.path="./figure/"}
qplot(x = interval, y = steps.avg, data = activity.per.5mins, geom = 'line')
```


### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps

```{r, echo=TRUE}
filter(activity.per.5mins, steps.avg == max(activity.per.5mins$steps.avg))
```


## Imputing missing values

### Number of missing values

```{r, echo=TRUE}
count(filter(activity, is.na(steps)))
```


### Filling missing values

```{r, echo=TRUE}
activity.filled <- activity %>% 
  inner_join(activity.per.5mins, by="interval") %>%
  mutate(steps.filled = ifelse(is.na(steps), steps.avg, steps)) %>%
  transmute(date = date, interval = interval, steps = steps.filled)
```


### Calculate total number of steps taken per day    

```{r, echo=TRUE}
activity.per.day.filled <- activity.filled %>% group_by(date) %>% summarize (steps = sum(steps))
```


### Show the total number of steps per day with an histogram

```{r, echo=TRUE, fig.path="./figure/"}
qplot(date, data=activity.per.day.filled, weight=steps, geom="histogram", binwidth=1)
```

Using imputation seems to change a little bit the graph, but not that much. Given the fact that missing values were replaced by the mean for the corresponding 5 minutes period, it doesn't change the mean at all... There's just a little bit less holes in the histogram.
 
### Mean and median of the total number of steps taken per day

```{r, echo=TRUE}
mean(activity.per.day.filled$steps, na.rm = T)
median(activity.per.day.filled$steps, na.rm = T)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE, fig.path="./figure/"}
activity.filled$day.type <- 
  as.factor(ifelse(weekdays(activity.filled$date) %in% c("Saturday","Sunday"), 'weekend', 'weekday'))

activity.per.5mins <- activity.filled %>% 
  group_by(interval, day.type) %>% 
  summarize (steps.avg = round(mean(steps, na.rm = T)))

qplot(interval, steps.avg, data = activity.per.5mins, facets = . ~ day.type, geom='line')
```
